Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-07-24 Grassmannian_cells.tex
Section 4 edits.

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym}
\usepackage[margin=1in]{geometry}
\usepackage{extarrows}
\input xy
\xyoption{all}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}
\newtheorem{proposition}[theorem]{Proposition}

@@

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

\section{Introduction}
A cluster algebra is a subalgebra of a rational function field where generators are defined by an explicit recursion.
Although the generators of the cluster algebra are given as rational functions they actually turn out to be Laurent polynomials.

@@

\noindent Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K(n)_0=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. By $S_1$ and $S_2$ we denote the BGP-reflection functor corresponding to the sink and source of $K(n)$ respectively \cite{bgp}. Note that the quiver $K(n)$ itself does not change when applying the reflection functors, but the vertices $1$ and $2$ are exchanged. Let $$\mathcal S:=\begin{cases}S_1\text{ if }1 \text{ is a source }\\S_2\text{ if }$2$ \text{ is a source }\end{cases}$$

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
\begin{theorem}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m=\mathcal S^{m-1} P_1$ and $I_m=\mathcal S^{-m+1} I_1$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.

@@

  To see (2), we apply the functor $\Hom(-,P_m)$ to the sequence \eqref{eq:truncated preprojectives} to get an isomorphism $\Hom(P_m^k,P_m)\cong\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$.
  Observe that $\Hom(P_m^k,P_m)$ is spanned by the coordinate projections $\pi_i:P_m^k\to P_m$ for $1\le i\le k$ and each of these gives rise to a commutative diagram
  \[\xymatrix{ & 0\ar[d] & 0\ar[d] & & \\ & P_m^{k-1}\ar[d]\ar@{=}[r]& P_m^{k-1}\ar[d]^{[f_1\ \cdots\ \widehat{f_i}\ \cdots\ f_k]}& & \\
    0\ar[r]& P_m^k\ar[r]^{[f_1\ \cdots\ f_k]}\ar[d]^{\pi_i}\ar@{}[dr]|(.35){\ulcorner}& P_{m+1}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]\ar@{=}[d]& 0 \\
    0\ar[r]& P_m\ar[r]^(.35){f_i}\ar[d]& P_{m+1}^{\{f_1,\ldots,\widehat{f_i},\ldots,f_k\}}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]& 0\\
     & 0 & 0 & & }\]
  The bottom row of this diagram is the extension in $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ which can be naturally identified with the generator $f_i$ of $\kk f_1\oplus\cdots\oplus\kk f_k$.

@@

\begin{lemma}
  Let $f_1,\ldots,f_n\in\Hom(P_m,P_{m+1})$ be linearly independent morphisms and write $g:P_{m-1}\to P_m$ for the composition of the inclusion from \eqref{eq:AR sequence} with the projection $\pi_n:P_m^n\to P_m$ to the $n^{th}$ factor.  Then $P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\cong P_m^{\{g\}}$.
\end{lemma}
\begin{proof}
  We have the following commutative diagram
  \[\xymatrix{0\ar[r]& P_m^{n-1}\ar@{=}[d]\ar[r]& P_m^n\ar[r]^{\pi_n}\ar[d]^{[f_1\ \cdots\ f_n]}\ar@{}[dr]|(.65){\lrcorner}& P_m\ar[r]\ar[d]^{f_n}& 0 \\
  0\ar[r]& P_m^{n-1}\ar[r]^{[f_1\ \cdots\ f_{n-1}]}& P_{m+1}\ar[r]& P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\ar[r]& 0}\]
  Since the vertical morphism on the left is an equality, the right hand square is a pullback.  
  But this implies the kernels of the vertical morphisms coincide.  
  The middle vertical morphism fits into the exact sequence \eqref{eq:AR sequence} and so the kernel is $P_{m-1}$.
  It follows that there is a morphism $g:P_{m-1}\to P_m$ so that $P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\cong P_m^{\{g\}}$.
\end{proof}

>>>>

\documentclass{amsart}
\usepackage{amsmath,amssymb,latexsym,color}
\usepackage[bookmarks=true,colorlinks=true, linkcolor=blue, citecolor=cyan]{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{extarrows}
\input xy
\xyoption{all}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}

@@

\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decomposition of Rank 2 Quiver Grassmannians}

@@

\section{Introduction}
A cluster algebra is a subalgebra of a rational function field where generators are defined by an explicit recursion \cite{fomin-zelevinsky}.
Although the generators of the cluster algebra are given as rational functions they actually turn out to be Laurent polynomials.
These Laurent polynomial expressions for the cluster variables have been studied extensively and ultimately can be described using the representation theory of quivers \cite{caldero-chapoton,caldero-keller}.

@@

\noindent Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K_0(n)=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. 
Write $\Sigma_1$ and $\Sigma_2$ for the BGP-reflection functors of $K(n)$ \cite{bgp}. 
Note that (up to relabeling vertices) the quiver $K(n)$ itself does not change when applying the reflection functors and thus we view them as endofunctors acting on the category $\rep K(n)$ of finite-dimensional representations.%, but the vertices $1$ and $2$ are exchanged. 
%Let $$\Sigma:=\begin{cases}\Sigma_1\text{ if vertex }1 \text{ is a source;}\\\Sigma_2\text{ if vertex }$2$ \text{ is a source.}\end{cases}$$

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
\begin{theorem}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  These representations satisfy the following recursions using the reflection functors:
  \[P_1=S_1,\quad P_m=\Sigma_1 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
  for $m\ge2$.
\end{remark}

@@

  To see (2), we apply the functor $\Hom(-,P_m)$ to the sequence \eqref{eq:truncated preprojectives} to get an isomorphism $\Hom(P_m^k,P_m)\cong\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$.
  Observe that $\Hom(P_m^k,P_m)$ is spanned by the coordinate projections $\pi_i:P_m^k\to P_m$ for $1\le i\le k$ and each of these gives rise to a commutative diagram
  \[\xymatrix{ & 0\ar[d] & 0\ar[d] & & \\ & P_m^{k-1}\ar[d]\ar@{=}[r]& P_m^{k-1}\ar[d]^{[f_1\ \cdots\ \widehat{f_i}\ \cdots\ f_k]}& & \\
    0\ar[r]& P_m^k\ar[r]^{[f_1\ \cdots\ f_k]}\ar[d]^{\pi_i}\ar@{}[dr]|(.35){\ulcorner}& P_{m+1}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]\ar@{=}[d]& 0 \\
    0\ar[r]& P_m\ar[r]^(.35){f_i}\ar[d]& P_{m+1}^{\{f_1,\ldots,\widehat{f_i},\ldots,f_k\}}\ar[r]\ar[d]& P_{m+1}^{\{f_1,\ldots,f_k\}}\ar[r]& 0\\
     & 0 & 0 & & }\]
  where $\hat f_k$ above indicates that the function $f_k$ has been omitted.
  The bottom row of this diagram is the extension in $\Ext(P_{m+1}^{\{f_1,\ldots,f_k\}},P_m)$ which can be naturally identified with the generator $f_i$ of $\kk f_1\oplus\cdots\oplus\kk f_k$.

@@

\begin{lemma}
  Let $f_1,\ldots,f_n\in\Hom(P_m,P_{m+1})$ be linearly independent morphisms and write $g:P_{m-1}\to P_m$ for the composition of the inclusion from \eqref{eq:AR sequence} with the projection $\pi_n:P_m^n\to P_m$ to the $n^{th}$ factor.  Then $P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\cong P_m^{\{g\}}$.
\end{lemma}
\begin{proof}
  We have the following commutative diagram
  \[\xymatrix{0\ar[r]& P_m^{n-1}\ar@{=}[d]\ar[r]& P_m^n\ar[r]^{\pi_n}\ar[d]^{[f_1\ \cdots\ f_n]}\ar@{}[dr]|(.65){\lrcorner}& P_m\ar[r]\ar[d]^{f_n}& 0 \\
  0\ar[r]& P_m^{n-1}\ \ \ \ar[r]^{[f_1\ \cdots\ f_{n-1}]}&\ \ \ P_{m+1}\ar[r]& P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}\ar[r]& 0}\]
  Since the vertical morphism on the left is an equality, the right hand square is a pullback.  
  But this implies the kernels of the vertical morphisms coincide.  
  The middle vertical morphism fits into the exact sequence \eqref{eq:AR sequence} and so the kernel is $P_{m-1}$.
  It follows that the resulting morphism $g:P_{m-1}\to P_m$, which is the kernel of the right-hand vertical morphism, satisfies $P_m^{\{g\}}\cong P_{m+1}^{\{f_1,\ldots,f_{n-1}\}}$.
\end{proof}
